A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning

### 2016-07-30 paper.tex
more
finished: notes for each DMN node to a model sub-task

@@

@@

@@

@@

@@

@@

@@

\item
  dmPFC:
  \begin{itemize}
    \item
dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
\item
largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
might be top-down modulated by more dmPFC subserved higher reflective and hypothetical processing.
consistent with present functional decoding, neural activity in the dmPFC, rather than vmPFC, has been con- sistently interpreted to underlie inference, representation, and assessment of one’s own and others’ mental states in functional neuroimaging research

@@

@@

\paragraph{Example.}
Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
  \begin{equation}
    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
  \end{equation}
  where
  $Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)$ is a normalization constant called the \textit{partition function} of $U$.
  It's not hard to see that $U$ and the $P$ above form a conjugate pair.

@@

>>>>

The model reinterprets many known effects

@@

We focus not on normative models (that describe what the brain does), but on process models (that describe how the brain does it)

@@

hierarchically deep models of the world

sensory exchanges with the world

show how a formal approach can provide generic explanations for psycho- pathology that are physiologically grounded

quantitative (and parametric) characterisations offered by computational

d Biological systems are homeostatic (or allostatic), which
means that they minimize the dispersion (entropy) of their
interoceptive and exteroceptive states.
d Entropy is the average of surprise over time, which means
that biological systems minimize the surprise associated
with their sensory states at each point in time.
d In statistics, surprise is the negative logarithm of B

perception as hypothesis testing

Consider an organism that, in their present situation, is confronted by several choices of what to do next. Being able to accurately and richly mentally enact possible future states before making a decision would help to evaluate the desirability of different outcomes and also the planning processes needed to make them happen.

 the use of this constructive process goes far beyond simply predicting the future, to the general evaluation of fitness for purpose

@@

vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 #3750} individual. Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled. Thus, dorsal regions of rostral mPFC contain value representations in the frame of reference of a modeled (as opposed to an actuated) preference irrespective of whether this applies to another’s or to one’s own likely actions and goals. Likewise, ventral mPFC contains a representation of value in the frame of reference of an executed choice, even if this executed choice reflects one’s own or another’s preferences. dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand. Critically, we also show this activity. While the simplest interpretation of this effect is that the region is simulating one’s own, currently irrelevant, preferences an alternative possibility is that the activity is projec- ting one’s own values into the mind of the partner, while simu- lating the partner’s choices. In essence, estimating the extent to which my own values would influence my partner, if they were making the choice. This iterated reasoning would be.

@@

DMN function has been linked to the recollection of past episodes and the simulation of future episodes

Simony 2016
This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system

@@

\item
the vmPFC to monitoring others’ (Lotze et al., 2007) and one’s own (Lane et al., 1997; Phan et al., 2004) emotional responses, that is, other’s (external) emotional reac- tions and one’s own (visceral) arousal. Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971), in line with the vmPFC’s functional association with general cognition and reward processing.

@@

\item
In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007). By contrast, its dorsomedial neighbor (rostral

\item
NAC was proposed to subserve model-free reinforcement learning in humans
(O'Doherty 2015)

@@

\item
  dmPFC:
  \begin{itemize}
    \item
dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways

\item
A class of neurons in macaque dorsomedial frontal cortex, including the presupplementary area, responds preferentially to the actions that another macaque makes rather than their own actions (Yoshida et al., 2011)
Many of the same neurons differentiate between another macaque’s actions that were unsuccessful, because the circumstances changed, and actions that were unsuccessful because the other macaque did not make the best choice given the circum- stances. The macaques adjusted their own subsequent actions in different ways after these two types of obser- vations. There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)

\item
largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
might be top-down modulated by more dmPFC subserved higher reflective and hypothetical processing.
consistent with present functional decoding, neural activity in the dmPFC, rather than vmPFC, has been con- sistently interpreted to underlie inference, representation, and assessment of one’s own and others’ mental states in functional neuroimaging research

@@

\item
hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future

\item
They found that amnesic patients with bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Figure 1). Moreover, Hassabis et al. [21] identified a possible source for this deficit. Although patients were able to produce considerable detail when asked to imagine fictitious new experiences, their descriptions lacked spatial coherence and were instead fragmented and considerably less rich [21].

\item
 empirical evidence hinting at a two-process function of the hippocampus comes from structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al. 2006b). Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al. 2006b; Woollett & Maguire 2009).

\item
Schachter/Buckner:
The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking (see Ref. [63]). As

@@

\paragraph{Example.}
Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
  \begin{equation}
    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
  \end{equation}
  where

  \begin{equation}
    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
  \end{equation}
  is a normalization constant called the \textit{partition function} of $U$.
  It's not hard to see that $U$ and the $P$ above form a conjugate pair.

@@

Friston2014:

This perspective shifts away from the brain as a passive filter of sensations (or an elaborate stimulus– response link) towards a view of the brain as a statistical organ that generates hypotheses or fantasies that are tested against sensory evidence. In short, the brain is now considered a phantastic organ (from Greek phantastikos, the ability to create mental images). For many people, this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).

Predictive coding is not a normative or descriptive scheme, it is a process theory with a biologically plausible basis—there is now much circumstantial anatomical and physiological evidence for predictive coding in the brain.12–16 In this scheme, neuronal representations in higher levels of cortical hierarchies generate predictions of representations in lower levels. These top-down predictions are compared with representations at the lower level to form a prediction
error (associated with the activity of superficial pyramidal cells). This mismatch signal is passed back up the hierarchy, to update higher representations (associated with the activity of deep pyramidal cells).

A description of biological processes as minimising (variational) free energy—a statistical measure of the surprise or improbability of sensory data, under a model of how those data were generated.

predictive coding represents a biologically plausible scheme that enables the brain to update beliefs about the world with sensory samples

 individuals spend most of their time predicting the internal (proprioceptive) and external (exteroceptive) consequences of behaviour (both their own and that of others). To fully appreciate the bilateral nature of these predictions, inference can be considered in an embodied context. In this setting, perception can be understood as resolving exteroceptive prediction errors by selecting the predictions that best explain sensations. Conversely, behaviour suppresses pro- prioceptive prediction error by changing proprioceptive sensations. This suppression relies on classic reflexes, in which equilibrium points are set by descending proprioceptive predictions (figure 1).15 This process is called active inference

providing a rich dynamical repertoire that enables the brain to respond quickly to changing inputs

the formal constraints implicit in predictive coding mandate modulatory gain control for ascending prediction errors
